Manifold Clustering in the Setting of Generalized Random Dot Product Graphs

### 2023-05-24 sdss-2023/slides.Rmd
minor revisions

\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
}
\For {$k = 1, ..., K$} {
Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}

>>>>

\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{i}$ to $X_k$ by minimizing $\sum_{i \in C_k} \|x_{i} - g_k(t_{i})\|^2$.\;
}
\For {$i = 1, ..., n$} {
Assign $z_i \leftarrow \arg\min_k \|x_i - g_k(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}

### 2023-05-24 sdss-2023/slides.tex
minor changes
minor updates
minor revisions

\begin{frame}{Manifold Block Model}
\protect\hypertarget{manifold-block-model}{}
Let \(p, q \geq 0\), \(d = p + q \geq 1\), \(1 \leq r < d\),
\(K \geq 2\), and \(n > K\) be integers. Define manifolds
\(\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}\) for
\(\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}\)
each by continuous function \(g_k : [0, 1]^r \to \mathcal{X}\). Define
probability distribution \(F\) with support \([0, 1]^r\). Then the
following mixture model is a \emph{manifold block model}:

@@

\begin{frame}{\(K\)-Curves Clustering}
\protect\hypertarget{k-curves-clustering}{}
\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
}
\For {$k = 1, ..., K$} {
Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}
\end{frame}

@@

where \(\hat{x}_i\) are the embedding vectors of \(A\). Suppose that for
each community \(k\), we have labels for at least \(R + 1\) vertices.
Then as \(n \to \infty\), \(K\)-curves clustering outputs estimates such
that

>>>>

\begin{frame}{Manifold Block Model}
\protect\hypertarget{manifold-block-model}{}
Let \(p, q \geq 0\), \(d = p + q \geq 1\), \(1 \leq r < d\),
\(K \geq 2\), and \(n > K\) be integers. Define manifolds
\(\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}\) for
\(\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}\)
each by continuous function \(g_k : [0, 1]^r \to \mathcal{X}\). Define
probability distribution \(F\) with support \([0, 1]^r\). Then the
following mixture model is a \emph{manifold block model}:

@@

\begin{frame}{\(K\)-Curves Clustering}
\protect\hypertarget{k-curves-clustering}{}
\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{i}$ to $X_k$ by minimizing $\sum_{i \in C_k} \|x_{i} - g_k(t_{i})\|^2$.\;
}
\For {$i = 1, ..., n$} {
Assign $z_i \leftarrow \arg\min_k \|x_i - g_k(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}
\end{frame}

@@

where \(\hat{x}_i\) are the embedding vectors of \(A\). Suppose that for
each community \(k\), we have labels for at least \(R + 1\) vertices.
Then as \(n \to \infty\), \(K\)-curves clustering outputs estimators
such that
